---
title: "analysis_per_team_reg"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
chooseCRANmirror(graphics=FALSE, ind=62)
if (!require("rsample")) install.packages("rsample") 
if (!require("dplyr")) install.packages("dplyr") 
if (!require("rpart")) install.packages("rpart") 
if (!require("rpart.plot")) install.packages("rpart.plot") 
if (!require("ipred")) install.packages("ipred") 
if (!require("randomForest")) install.packages("randomForest")
if (!require("ranger")) install.packages("ranger") 
if (!require("caret")) install.packages("caret") 
if (!require("MASS")) install.packages("MASS") 
if (!require("e1071")) install.packages("e1071")

library(rsample)      # data splitting, functions: initial_split 
library(dplyr)        # data wrangling
library(rpart)        # performing regression trees
library(rpart.plot)   # plotting regression trees
library(ipred)        # bagging
library(randomForest) # basic implementation
library(ranger)       # a faster implementation of randomForest
library(caret)        # an aggregator package for performing many machine learning models
library(MASS)         # dataset
library(e1071)
```


Load in the data
```{r}
setwd("C:/Users/senic/OneDrive/Desktop/Masters/FALL_2021/ECO1400/Term_Paper/Data")
dataset_full <- read.csv(file="All_Disper_1_2_3_2011_2018.csv", header=TRUE, sep=",", na = ".")
names(dataset_full)
```
```{r}
dataset_full <- dataset_full[dataset_full$HomeAveSal != 0, ]
dataset_full <- dataset_full[dataset_full$HomeAveSal != 0, ]
```

```{r}
summary(dataset_full$HomeScore)
summary(dataset_full$AwayScore)

```


Create training (70%) and test (30%) sets for the datasetHousing::make_dataset() data.
Use set.seed for reproducibility
```{r}
na.omit(dataset_full)

set.seed(123)
dataset_split <- initial_split(dataset_full, prop = 0.7)
dataset_train <- training(dataset_split)
dataset_test  <- testing(dataset_split)
```


Fit a regression tree
```{r}
m1 <- rpart(formula = HomeScore ~ HomeMaxSal + HomeAveSal + HomeDisper1 + HomeDisper2 + HomeDisper3, data = dataset_train, method = "anova")
```


Visualize the fitted tree 
```{r}
rpart.plot(m1, cex=0.5)
```


*************************** RANDOM FORESTS ***************************

Random forests are a modification of bagging that builds a large collection of de-correlated trees, by inducing more randomness into the tree growing process. This is achieved in two stages:

1. Bootstrap: similar to bagging, where each tree is grown to a bootstrap resampled data set.
2. Split-variable randomization: each time a split is to be performed, the search for the split variable is limited to a random subset of m of the p variables. 

Fit a random forest model
```{r}
rfHome <- randomForest(formula = HomeScore ~ HomeMaxSal + HomeAveSal + HomeDisper1 + HomeDisper2 + HomeDisper3, data = dataset_train, ntrees = 100, na.action=na.omit, importance=TRUE)
rfHome
rfAway <- randomForest(formula = AwayScore ~ AwayMaxSal + AwayAveSal + AwayDisper1 + AwayDisper2 + AwayDisper3, data = dataset_train, ntrees = 100, na.action=na.omit, importance=TRUE)
rfAway
rfHomeFull <- randomForest(formula = HomeScore ~ HomeMaxSal + AwayMaxSal + HomeAveSal + AwayAveSal + HomeDisper1 + AwayDisper1 + HomeDisper2 + AwayDisper2 + HomeDisper3 + AwayDisper3, data = dataset_train, ntrees = 100, na.action=na.omit, importance=TRUE)
rfHomeFull
rfAwayFull <- randomForest(formula = AwayScore ~ HomeMaxSal + AwayMaxSal + HomeAveSal + AwayAveSal + HomeDisper1 + AwayDisper1 + HomeDisper2 + AwayDisper2 + HomeDisper3 + AwayDisper3, data = dataset_train, ntrees = 100, na.action=na.omit, importance=TRUE)
rfAwayFull

```



Plot the error rate
```{r, fig.width=4, fig.height=3}
plot(rfHome)
plot(rfAway)
plot(rfHomeFull)
plot(rfAwayFull)
```

Out of sample prediction
```{r}
pred_randomForestHome <- predict(rfHome, newdata = dataset_test)
head(pred_randomForestHome)
pred_randomForestAway <- predict(rfAway, newdata = dataset_test)
head(pred_randomForestAway)
pred_randomForestHomeFull <- predict(rfHomeFull, newdata = dataset_test)
head(pred_randomForestAway)
pred_randomForestAwayFull <- predict(rfAwayFull, newdata = dataset_test)
head(pred_randomForestAway)
```

Assess out-of-sample predictive accuracy
```{r}
RMSEHome <- RMSE(pred_randomForestHome, dataset_test$HomeScore)
RMSEHome
RMSEAway <- RMSE(pred_randomForestAway, dataset_test$AwayScore)
RMSEAway
RMSEHomeFull <- RMSE(pred_randomForestHomeFull, dataset_test$AwayScore)
RMSEHomeFull
RMSEAwayFull <- RMSE(pred_randomForestAwayFull, dataset_test$AwayScore)
RMSEAwayFull
```

Compare to OLS
```{r}
base1 <- lm(formula = HomeScore ~ HomeMaxSal + HomeAveSal + HomeDisper1 + HomeDisper2 + HomeDisper3, dataset_full)
yhat_base1 <- predict(base1, newdata = as.data.frame(dataset_test))
base1_resid <- dataset_test$deltaScore - yhat_base1
RMSE_base1 <- sqrt(mean(base1$residuals^2))
RMSE_base1

base2 <- lm(formula = AwayScore ~ AwayMaxSal + AwayAveSal + AwayDisper1 + AwayDisper2 + AwayDisper3, dataset_full)
yhat_base2 <- predict(base2, newdata = as.data.frame(dataset_test))
base2_resid <- dataset_test$deltaScore - yhat_base2
RMSE_base2 <- sqrt(mean(base2$residuals^2))
RMSE_base2

base3 <- lm(formula = HomeScore ~ HomeMaxSal + AwayMaxSal + HomeAveSal + AwayAveSal + HomeDisper1 + AwayDisper1 + HomeDisper2 + AwayDisper2 + HomeDisper3 + AwayDisper3, dataset_full)
yhat_base3 <- predict(base3, newdata = as.data.frame(dataset_test))
base3_resid <- dataset_test$deltaScore - yhat_base3
RMSE_base3 <- sqrt(mean(base3$residuals^2))
RMSE_base3

base4 <- lm(formula = AwayScore ~ HomeMaxSal + AwayMaxSal + HomeAveSal + AwayAveSal + HomeDisper1 + AwayDisper1 + HomeDisper2 + AwayDisper2 + HomeDisper3 + AwayDisper3, dataset_full)
yhat_base4 <- predict(base4, newdata = as.data.frame(dataset_test))
base4_resid <- dataset_test$deltaScore - yhat_base4
RMSE_base4 <- sqrt(mean(base4$residuals^2))
RMSE_base4
```

```{r}
(RMSE_base1 - RMSEHome)/RMSE_base1 * 100
(RMSE_base2 - RMSEAway)/RMSE_base2 * 100
(RMSE_base3 - RMSEHomeFull)/RMSE_base3 * 100
(RMSE_base4 - RMSEAwayFull)/RMSE_base4 * 100
```






```{r}
varImpPlot(rfHome)
varImpPlot(rfAway)
varImpPlot(rfHomeFull)
varImpPlot(rfAwayFull)
```










